---
title: "R Notebook Richness figure creation"
output: html_notebook
---

## Load library

```{r}

library(grid)
library(cowplot)
library(ggplot2)


```




```{r}

Grp_color = c("#dfc27d","#a6bddb","#74a9cf","#3690c0","#0570b0","#045a8d","#023858")


```


## Figure 1

Figure 1: 
Panel A: model de courbes de models alternatives
Panel B: study design with sampling design (julien)
Panel C: microbiota shifts with diet (Richess OTU)
Panel D: DSS – dose response (colong length, mono, globlet, poids?)
Panel E: histo – Pictures

```{r, fig.height=12, fig.width=12}

study_design <- png::readPNG("study_design_bis.png")
hypothesis   <- png::readPNG("hypothesis_model.png")
histology    <- png::readPNG("histology_DSS.png")

study_design_grob1 <- rasterGrob(study_design)
hypothesis_grob2   <- rasterGrob(hypothesis)
histology_grob3    <- rasterGrob(histology)


load("diet_press_plot.rda")
load("p_physio.rda")

cowplot::plot_grid(plot_grid(hypothesis_grob2, study_design_grob1, rel_widths = c(1,2), labels=c("A","B")), plot_grid(pB_diet+ scale_color_manual(values=Grp_color[2:1]), histology_grob3,labels=c("C","D"), rel_widths = c(1.2,1)), p_physio, ncol=1, rel_heights = c(1,1,1.5), labels=c("","","E"))

ggsave("Figure1.pdf",h=12,w=12)



```


## Supplementary Figure 1

- Data on the impact of diet on microbiota (species/ composition) =>  faire apparaitre les OTU qui disparaissent (poissblement dans la légende de la figure)
- Enlever le point -50 (@Julien)


```{r, fig.height=5, fig.width=12, message=FALSE, warning=FALSE}

cowplot::plot_grid(
pA_diet + scale_color_manual(values=Grp_color[1:2]) + guides(color=FALSE),
pC_diet + scale_color_manual(values=Grp_color[1:2]), labels="AUTO", rel_widths = c(1.1,2)
)

ggsave("Supp_Figure1.pdf",h=5,w=12)


#most of OTU disaparing fall into discriminating genus


```








