---
title: "Richness R Notebook : DSS press perturbation"
output:
  html_notebook: 
    highlight: tango
    author: Julien Tap
    mathjax: null
    theme: cerulean
    toc: yes
    toc_float: yes
    code_folding: hide
  html_document: default
---



```{r}

Grp_color = c("#dfc27d","#a6bddb","#74a9cf","#3690c0","#0570b0","#045a8d","#023858")


```





## Loading 16S sequencing data and associated metadata

here we start to load all R libraries and custom script (code not shown) with study data

```{r loading_library2, message=FALSE, warning=FALSE, include=FALSE}

Sys.setenv(PATH = paste("C:/Users/tapju/Documents/ImageMagick", Sys.getenv("PATH"), sep = ";"))
library(BiotypeR)
library(DirichletMultinomial)
library(ggplot2)
library(reshape2)
library(ade4)
library(vegan)
library(dplyr)
library(magrittr)
library(forcats)
library(GGally)
source("https://raw.githubusercontent.com/tapj/IBSMicrobiota/master/R/mclapply.hack.R")
library(stringr)
library(broom)

library(earlywarnings)
library(gganimate)
library(cowplot)
library(microbiome)
library(viridis)
source("PlotPotential2.r")

```

```{r}

rat_metadata = read.csv2("rat_metadata.csv")
#mapping_file = read.table("qiime/result/Mapping_Richness1.txt", header=T, comment.char ="", dec=".", sep="\t")

mapping_file = read.table("qiime/result/mapping_all_richness.txt", header=T, comment.char ="", dec=".", sep="\t")

#rat_metadata$Animal_id_2 %<>% gsub("R","",.)

genus = read.table("qiime/result/otus/summarize_taxa_through_plots_out/otu_table_non_chimeric_L6.txt", header=T, comment.char ="", dec=".", sep="\t", skip=1, row.names = 1, check.names = FALSE)

otu   = read.table("qiime/result/otus/make_otu_table_out/otu_table_non_chimeric.tab",header=T, comment.char ="", dec=".", sep="\t", skip=1, row.names = 1, check.names = FALSE )

colnames(mapping_file)[7] = "Time"


mapping_file

```





## select sample of interest


```{r}




mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  pull(samplenr) -> select_DSS_samples
  




```


## alpha-diversity


```{r}
observed_otu = otu[,-dim(otu)[2]] %>% t %>% vegan::rarefy(.,sample=25000)


observed_otu %<>% as.matrix %>% as.data.frame %>% rename(S_obs = V1 )

mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  merge(.,observed_otu, by.x="samplenr", by.y="row.names", all = FALSE) %>%
  as_tibble %>%
  mutate(Time = Time %>% as.character) %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
   #xtabs(~Time+Study_group, data = .) %>%
  ggplot() + 
  #geom_line(aes(x=DSS,y=S_obs, group=Rat_ID), linetype=2) +
  geom_point(aes(x=DSS,y=S_obs), size=1, alpha=0.5) +
  xlab("DSS") + ylab("Observed OTU\nafter reads rarefaction") +
  facet_wrap(~Time_days)


```

```{r, fig.height=4, fig.width=10}

mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  merge(.,observed_otu, by.x="samplenr", by.y="row.names", all = FALSE) %>%
  as_tibble %>%
  mutate(Time = Time %>% as.character) %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  mutate(DSS_group = ifelse(DSS>1,"High dose","No or Low dose")) %>%
  group_by(DSS_group,Time_days) %>%
  summarise(`Standard deviation`=sd(S_obs),Average=median(S_obs)) %>%
  melt(id.vars = c("DSS_group","Time_days")) %>%
   #xtabs(~Time+Study_group, data = .) %>%
  ggplot() + 
  #geom_line(aes(x=DSS,y=S_obs, group=Rat_ID), linetype=2) +
  geom_line(aes(x=Time_days,y=value, col=DSS_group), size=2, alpha=0.5, stat = "identity",linetype=1) +
  geom_point(aes(x=Time_days,y=value, col=DSS_group), size=4) +
  xlab("Time (days)") + ylab("Observed OTU\nstatistics") + facet_wrap(~variable, scale="free_y") -> pD_Sobs

pD_Sobs

```



## beta-diversity


```{r}


genus_jsd = genus[, select_DSS_samples %>% as.character] %>%
  dist.JSD()

genus_bc =  genus %>% BiotypeR::noise.removal(percent=0.1) %>% t %>% vegan::vegdist() %>% as.matrix %>% as.dist


genus_pco = dudi.pco(genus_jsd, scannf=F, nf=3)


genus_pco_bc = dudi.pco(genus_bc, scannf=F, nf=3)


#TO DO : distance vs baseline plot

```


### distance vs baseline
```{r, fig.height=6, fig.width=12, message=FALSE, warning=FALSE}




mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -7 ) %>% 
  filter(Time != "T47") %>%
  merge(.,genus_pco$li, by.x="samplenr", by.y="row.names") -> selected_metadata_DSS

plot(1:10)

genus_jsd %>%
  as.matrix() %>%
  reshape2::melt() %>%
  merge(.,selected_metadata_DSS %>% select(-A1,-A2,-A3), by.x="Var1", by.y="samplenr") %>%
  merge(.,selected_metadata_DSS %>% select(-A1,-A2,-A3), by.x="Var2", by.y="samplenr") %>%
  filter(Rat_ID.x == Rat_ID.y ) %>%
  mutate(Step  =  
           
           case_when(
             Time_days.x == -7 & Time_days.y == -1 ~  "Water",
             Time_days.x == -1 & Time_days.y == 3 ~  "DSS",
             Time_days.x == 13 & Time_days.y == 17 ~  "DSS",
             Time_days.x == 27 & Time_days.y == 31 ~  "DSS",
             Time_days.x == 3 & Time_days.y == 13 ~  "Water",
             Time_days.x == 17 & Time_days.y == 27 ~  "Water",
             Time_days.x == 31 & Time_days.y == 40 ~  "Water",
             Time_days.x == 40 & Time_days.y == 68 ~  "Water")
         ) %>%
  mutate(days  =  
           
           case_when(
             Time_days.x == -7 & Time_days.y == -1 ~  0,
             Time_days.x == -1 & Time_days.y == 3 ~  3,
             Time_days.x == 13 & Time_days.y == 17 ~  17,
             Time_days.x == 27 & Time_days.y == 31 ~  31,
             Time_days.x == 3 & Time_days.y == 13 ~  13,
             Time_days.x == 17 & Time_days.y == 27 ~  27,
             Time_days.x == 31 & Time_days.y == 40 ~  40,
             Time_days.x == 40 & Time_days.y == 68 ~  68)
         ) %>%
  filter(!is.na(Step)) %>%
  filter(Study_group.y %in% c("G2","G3","G4","G5","G6","G7")) %>%
   group_by(Study_group.y,Step,days,Time_days.x, Time_days.y) %>%
  summarize(m=1-median(value), q25=1-quantile(value,0.25), q75=1-quantile(value,0.75) ) %>%
  
  ggplot() + 
  geom_line(aes(x=days  , y=m, group=Study_group.y, col=Study_group.y), size=3) +
  geom_line(aes(x=days  , y=q25, group=Study_group.y, col=Study_group.y), linetype=2  ) +
  geom_line(aes(x=days  , y=q75, group=Study_group.y, col=Study_group.y), linetype=2) +
 
  
  geom_hline(yintercept = 1-0.19, col="grey", alpha=0.5, size=3) +
  geom_rect(aes(xmin=Time_days.x,xmax=Time_days.y, fill=Step), alpha=0.2, ymin=-Inf, ymax=+Inf) +
  
  #geom_vline(xintercept = c(1:8) ,  col=rep(c("blue","red","blue","red","blue","red","blue","blue"),6),  alpha=0.1, size=10) +
   facet_wrap(~Study_group.y) + scale_fill_manual(values=c("Red","Blue")) + 
  guides(col=FALSE) + ylab("Microbiota Stability\n(1-JSD per step)") + xlab("Study steps")
  
  
ggsave("Figure_jsd_stability_DSS.pdf", h=6, w=12)




genus_jsd %>%
  as.matrix() %>%
  reshape2::melt() %>%
  merge(.,selected_metadata_DSS %>% select(-A1,-A2,-A3), by.x="Var1", by.y="samplenr") %>%
  merge(.,selected_metadata_DSS %>% select(-A1,-A2,-A3), by.x="Var2", by.y="samplenr") %>%
  filter(Rat_ID.x == Rat_ID.y ) %>%
  mutate(Step  =  
           
           case_when(
             Time_days.x == -7 & Time_days.y == -1 ~  "Water",
             Time_days.x == -1 & Time_days.y == 3 ~  "DSS",
             Time_days.x == 13 & Time_days.y == 17 ~  "DSS",
             Time_days.x == 27 & Time_days.y == 31 ~  "DSS",
             Time_days.x == 3 & Time_days.y == 13 ~  "Water",
             Time_days.x == 17 & Time_days.y == 27 ~  "Water",
             Time_days.x == 31 & Time_days.y == 40 ~  "Water",
             Time_days.x == 40 & Time_days.y == 68 ~  "Water")
         ) %>%
  mutate(days  =  
           
           case_when(
             Time_days.x == -7 & Time_days.y == -1 ~  0,
             Time_days.x == -1 & Time_days.y == 3 ~  3,
             Time_days.x == 13 & Time_days.y == 17 ~  17,
             Time_days.x == 27 & Time_days.y == 31 ~  31,
             Time_days.x == 3 & Time_days.y == 13 ~  13,
             Time_days.x == 17 & Time_days.y == 27 ~  27,
             Time_days.x == 31 & Time_days.y == 40 ~  40,
             Time_days.x == 40 & Time_days.y == 68 ~  68)
         ) %>%
  filter(!is.na(Step)) %>%
  merge(.,hysteresis, by.x="Rat_ID.x", by.y="Rat_ID") %>%
  filter(Study_group.y %in% c("G2","G5","G6","G7")) %>%
  mutate(hysteresis_grp = ifelse(Study_group.y == "G2","control (DSS 0%)",hysteresis_grp %>% as.character)) %>%
  group_by(hysteresis_grp,Step,days,Time_days.x, Time_days.y) %>%
  summarize(m=1-median(value), q25=1-quantile(value,0.25), q75=1-quantile(value,0.75) ) %>%
  
  ggplot() + 
  geom_line(aes(x=days  , y=m, group=hysteresis_grp, col=hysteresis_grp), size=3) +
  geom_line(aes(x=days  , y=q25, group=hysteresis_grp, col=hysteresis_grp), linetype=2  ) +
  geom_line(aes(x=days  , y=q75, group=hysteresis_grp, col=hysteresis_grp), linetype=2) +
 
  geom_hline(yintercept = 1-0.19, col="grey", alpha=0.5, size=3) +
  #geom_vline(aes(xintercept = days-1) ,  col=rep(c("blue","red","blue","red","blue","red","blue","blue"),4),  alpha=0.1, size=3) +
  geom_rect(aes(xmin=Time_days.x,xmax=Time_days.y, fill=Step), alpha=0.2, ymin=-Inf, ymax=+Inf) +
   #theme(axis.text.x = element_text(hjust =0.5)) +
   facet_wrap(~hysteresis_grp) +
  scale_fill_manual(values=c("Red","Blue")) +
  guides(col=FALSE) + ylab("Microbiota Stability\n(1- JSD per step)") + xlab("Days") 

ggsave("Figure_jsd_stability_resilience.pdf", h=6, w=12)






```




```{r, fig.height=8, fig.width=5}

mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  merge(.,genus_pco$li, by.x="samplenr", by.y="row.names") %>%
  ggplot() + geom_point(aes(x=A1,y=A2,col=Study_group)) + 
  scale_color_brewer() +
  facet_wrap(~Time_days, ncol=2) +
  theme_dark()
  
  
mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  merge(., genus_pco_bc$li, by.x="samplenr", by.y="row.names") %>%
  ggplot() + geom_point(aes(x=A1,y=A2,col=Study_group)) + 
  scale_color_brewer() +
  facet_wrap(~Time_days, ncol=2) +
  theme_dark()
  
  genus_jsd_melt = genus_jsd %>% broom::tidy()
  
```



```{r}
  
  
    
mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  merge(genus_jsd %>% as.matrix %>% melt(), by.x="samplenr", by.y="Var1") %>%
  merge(., mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") , by.x="Var2", by.y="samplenr") %>%
  filter(Time_days.x == -7, Rat_ID.x == Rat_ID.y) %>%
  ggplot() + geom_boxplot(aes(x=Time_days.y %>% as.character, y=value)) + facet_wrap(~Study_group.y)




```


```{r}


mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  merge(.,genus_pco$li, by.x="samplenr", by.y="row.names") %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  select(contains("Akkermansia"),contains("S24-7"),DSS,Time_days, Rat_ID,A1,A2,A3) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","Low dose")) %>%
  mutate(Pertubation = ifelse(Time_days>=31,"DSS pulse","washout")) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"S24_7") %>%
  ggplot() + geom_point(aes(x=A1,y=A2,col=log10(Akkermansia+10^-6))) + 
  #scale_color_brewer() +
  viridis::scale_color_viridis("Akkermansia",option = "B") +
  #facet_grid(Pertubation~DSS_group) +
  theme_dark() + xlab("PC1") + ylab("PC2") -> pA_akk
  
  pA_akk

```

### plot akk and PCoA
```{r}

PD_akk_PCoA_plot =

  mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  merge(.,genus_pco$li, by.x="samplenr", by.y="row.names") %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  select(contains("Akkermansia"),contains("S24-7"),DSS,Time_days, Rat_ID,A1,A2,A3) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","Low dose")) %>%
  mutate(Pertubation = ifelse(Time_days>=31,"DSS pulse","washout")) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"S24_7") %>%
  ggplot() + geom_point(aes(x=A1,y=(Akkermansia+10^-6))) + #theme_dark() + 
  xlab("PCo1 (JSD based)") + ylab("Akkermansia\nrel. abund.") + scale_y_continuous(labels=scales::percent)


PD_akk_PCoA_plot
```




```{r, fig.height=8, fig.width=5}


mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,genus_pco$li, by.x="samplenr", by.y="row.names") %>%
  ggplot() + geom_point(aes(x=DSS,y=A2)) + 
  scale_color_brewer() +
  facet_wrap(~Time_days, ncol=2) +
  geom_smooth(aes(x=DSS+0.125,y=A2)) +
  scale_x_continuous(trans="log2") +
  theme_dark()




```


```{r}



mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,genus_pco$li, by.x="samplenr", by.y="row.names") %>%
  ggplot() + geom_point(aes(x=Time_days,y=A3)) + 
  scale_color_brewer() +
  facet_wrap(~DSS, ncol=2) +
  #geom_smooth(aes(x=DSS+0.125,y=A2)) +
  #scale_x_continuous(trans="log2") +
  theme_dark()






```




```{r, message=FALSE, warning=FALSE}


genus_stool.int.denoized =  round(noise.removal(genus[, select_DSS_samples %>% as.character], percent=1) * 30000)



fit_genus_DSS_list = vector("list",5)

if(!file.exists("fit_genus_DSS_list.rda")) { #check if fit_genus_list.rda exist
#fit_genus <- mclapply(1:6, dmn, count=t(genus.count), verbose=TRUE)

set.seed(444); seeds=sample(1:1000, 5)

  for(i in 1:5) {
    set.seed(seeds[i])
    fit_genus <- mclapply(1:15, dmn, count=t(genus_stool.int.denoized), verbose=TRUE)
    fit_genus_DSS_list[[i]] = fit_genus
  }
  #save(fit_genus, file=("../inst/extData/fit_genus.rda"))
  save(fit_genus_list, file=("fit_genus_DSS_list.rda"))

} else {load("fit_genus_DSS_list.rda")}





lplc = vector("list",5)

for(i in 1:5) {
  lplc[[i]] <- sapply(fit_genus_DSS_list[[i]], function(x){attr(x,"goodnessOfFit")[["Laplace"]]})
}

(best_genus <- fit_genus_DSS_list[[5]][[which.min(lplc[[1]])]])
#best_genus <- fit_genus_DSS_list[[1]][[11]] # true one

#best_genus <- fit_genus_DSS_list[[1]][[6]] # ideal one


#bic_results=melt(data.frame(sapply(fit_genus_DSS_list[1:5], sapply, BIC), nb.clust=1:15), id.vars="nb.clust")
laplace_results=melt(data.frame(sapply(fit_genus_DSS_list[1:5], sapply, laplace), nb.clust=1:15), id.vars="nb.clust")

#p1_bic = ggplot(bic_results) + geom_line(aes(x=nb.clust, y=value, group=variable)) + ylab("BIC") + xlab("Number of clusters")
p2_laplace = ggplot(laplace_results) + geom_line(aes(x=nb.clust, y=value, group=variable)) + ylab("Laplace") + xlab("Number of clusters")

#p1_bic + theme_bw() 
p2_laplace + theme_bw() # to complete this task but K = 5

par(mar=c(1,1,1,10))
DirichletMultinomial::heatmapdmn(fit1=fit_genus_DSS_list[[5]][[1]], fitN=best_genus, count=t(genus_stool.int.denoized))





```



```{r}



genus_drivers = 
rownames(genus_stool.int.denoized)[rev(head(order(  rowSums(abs(fitted(fit_genus[[9]], scale = TRUE) - as.vector(fitted(fit_genus[[1]], scale = TRUE))))  , decreasing = TRUE), 10))]



 f = fitted(fit_genus[[9]], scale = TRUE) - as.vector(fitted(fit_genus[[1]], scale = TRUE))
 
 f %>% 
   abs %>%
  apply(1,sum) %>%
   as_tibble() %>%
   tibble::rownames_to_column("genus") %>%
    arrange(desc(value)) %>%
  top_n(10) %>%
   tidyr::separate(genus,into=c("Domain","Phylum","Class","Order","Family","Genus"),sep=";") %>%
   ggplot() + geom_bar(aes(y=value, x=paste0("[",Family %>% gsub("D_4__","",.),"] ",Genus %>% gsub("D_5__","",.)) %>% reorder(.,value) ), stat="identity") +
   coord_flip() + xlab("") + ylab("Genus DMM weights") -> pB_akk


pB_akk


```

### bistability test

Bistable genus analysis on Dominant Genera which contribute to the DMM models
```{r}

b = apply(
  log10(genus[genus_drivers,select_DSS_samples %>% as.character]+10^-4), 1, function(x) 1-microbiome::bimodality(x,bs.iter = 100) 
  ) %>% round(3) 

data.frame(tax = b %>% names, p.values = b) %>% 
  arrange(p.values) #%>%   filter(p.values <= 0.05)


bistable_dominant_genera = 
  data.frame(tax = b %>% names, p.values = b) %>% 
  arrange(p.values) %>% 
  filter(p.values < 0.05) %>%
  .$tax %>% as.character

tipping_point = 
  apply(log10(genus[bistable_dominant_genera,]+10^-4), 1,   function(x) microbiome::potential_analysis(x,bs.iter = 100)$minima) %>%
  lapply(.,function(x) 10^x ) %>% 
  lapply(.,max) %>% 
  as.data.frame(check.names=FALSE) %>% t %>% 
  as.data.frame(check.names=FALSE) %>% set_colnames("Tip")

merge(t(genus[bistable_dominant_genera,])+10^-4, mapping_file[,c("X.SampleID", "Time", "DSS")], by.x="row.names", by.y="X.SampleID") %>%
  melt(.,id.vars=c("Row.names","Time", "DSS")) %>% 
  merge(.,tipping_point, by.x="variable", by.y="row.names", all = TRUE ) %>%
  filter(DSS == "3" | DSS == "2" | DSS == "1") %>%
  ggplot() + 
  geom_histogram(aes(x=value), fill="white", bins = 20) + geom_density(aes(x=value, y = ..count../6)) +
  facet_wrap(~gsub(";","\n",gsub("D_","",variable)), scale="free") + 
  scale_x_log10() + theme_dark() + 
  geom_vline(aes(xintercept = Tip), col="red", size=2, alpha = 0.75) + scale_y_log10() #+ geom_density(aes(x=value))

tipping_point 



```



```{r}


library(ggridges)


mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),contains("D_5__Ruminococcus"),contains("Phascolarctobacterium"),contains("D_4__Christensenellaceae;D_5__uncultured"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"Ruminococcus") %>%
  rename_at(3, ~"Phascolarctobacterium") %>%
  rename_at(4, ~"Christensenellaceae") %>%
ggplot() + #geom_point(aes(x=Akkermansia,y=Time_days))
  geom_density_ridges_gradient(aes(x=Akkermansia + 10^-6, y=reorder(Time_days %>% as.character,Time_days), fill=..x..)) + 
  #geom_density(aes(x=Bifidobacterium + 10^-6, fill=cat_age_val)) +
  
  facet_wrap(~DSS_group, scales ="free_y", ncol=4) +
  viridis::scale_fill_viridis(option = "B") +
  geom_vline(xintercept = 0.009129289, col="red", size=1, alpha = 0.75, linetype=3) +
  guides(fill=FALSE) +
  xlab("Akkermansia") + ylab("stool sample prob over time") +
  scale_x_log10() + theme_dark() -> pC_Akk


pC_Akk


```


```{r, fig.height=10, fig.width=6}



library(ggridges)


mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),contains("D_5__Ruminococcus"),contains("Phascolarctobacterium"),contains("D_4__Christensenellaceae;D_5__uncultured"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"Ruminococcus") %>%
  rename_at(3, ~"Phascolarctobacterium") %>%
  rename_at(4, ~"Christensenellaceae") %>%
  reshape2::melt(id.vars=c("DSS","Time_days","Rat_ID","DSS_group")) %>%
ggplot() + #geom_point(aes(x=Akkermansia,y=Time_days))
  geom_density_ridges_gradient(aes(x=value + 10^-6, y=reorder(Time_days %>% as.character,Time_days), fill=..x..)) + 
  #geom_density(aes(x=Bifidobacterium + 10^-6, fill=cat_age_val)) +
  
  facet_grid(variable~DSS_group, scales ="free") +
  viridis::scale_fill_viridis(option = "B") +
  #geom_vline(xintercept = 0.009129289, col="red", size=1, alpha = 0.75, linetype=3) +
  guides(fill=FALSE) +
  xlab("rel. abundance") + ylab("stool sample prob over time") +
  scale_x_log10() + theme_dark() -> p_all_bimodal_species


p_all_bimodal_species

ggsave(filename = "Figure_over_time_bimodal_species.pdf")


```




```{r}


library(ggridges)


mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -8) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),contains("D_5__Ruminococcus"),contains("Phascolarctobacterium"),contains("D_4__Christensenellaceae;D_5__uncultured"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose (DSS >= 1%)","No or Low dose (DSS < 1%)")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"Ruminococcus") %>%
  rename_at(3, ~"Phascolarctobacterium") %>%
  rename_at(4, ~"Christensenellaceae") %>%
ggplot() + #geom_point(aes(x=Akkermansia,y=Time_days))
  geom_density_ridges_gradient(aes(x=Akkermansia + 10^-6, y=reorder(Time_days %>% as.character,-Time_days), fill=..x..)) + 
  #geom_density(aes(x=Bifidobacterium + 10^-6, fill=cat_age_val)) +
  
  facet_wrap(~DSS_group, scales ="free_y", ncol=4) +
  
  geom_vline(xintercept = 0.009129289, col="red", size=1, alpha = 0.75, linetype=3) +
  guides(fill=FALSE) +
  xlab("Akkermansia") + ylab("stool sample prob over time") +
  scale_fill_viridis(option = "A") +
  scale_x_log10() + theme_dark() -> pC_Akk


pC_Akk




```



```{r, fig.height=8, fig.width=14}


  plot_grid(pD_Sobs,pB_akk,pA_akk,pC_Akk, ncol=2,labels = c("A","B","C","D"), rel_heights = c(2,2))

ggsave("Figure_early_warnings_signal.pdf")





```


# Akkermansia Time serie pairwise comparision

```{r}


mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),contains("D_5__Ruminococcus"),contains("Phascolarctobacterium"),contains("D_4__Christensenellaceae;D_5__uncultured"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"Ruminococcus") %>%
  rename_at(3, ~"Phascolarctobacterium") %>%
  rename_at(4, ~"Christensenellaceae") %>%
  filter(DSS_group == "High dose") %>%
  select(Akkermansia,Rat_ID,Time_days) %>%
  #mutate(Akkermansia = log10(Akkermansia+10^-6)) %>%
  dcast(Time_days~Rat_ID, value.var="Akkermansia", fill=NA) %>%
  #.[c(1:9,7:9),] %>%
  select(-Time_days) %>%
  t %>%
  fpc::pamk( metric="euclidean") -> akk_pamk
  




  akk_pamk$pamobject$clustering
  

  
  



```



```{r, fig.height=4, fig.width=10}

mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),contains("D_5__Ruminococcus"),contains("Phascolarctobacterium"),contains("D_4__Christensenellaceae;D_5__uncultured"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  rename_at(2, ~"Ruminococcus") %>%
  rename_at(3, ~"Phascolarctobacterium") %>%
  rename_at(4, ~"Christensenellaceae") %>%
  filter(DSS_group == "High dose") %>%
  select(Akkermansia,Rat_ID,Time_days) %>%
  merge(., akk_pamk$pamobject$clustering %>% as.data.frame %>% rename_at(1, ~"cluster"), by.x="Rat_ID", by.y="row.names") %>%
  ggplot() + 
  geom_line(aes(x=Time_days,y=Akkermansia + 10^-6, group=Rat_ID), linetype=3) + 
  geom_point(aes(x=Time_days,y=Akkermansia + 10^-6,  col=cluster %>% as.character)) +
  scale_y_log10() + facet_wrap(~cluster) + 
  geom_smooth(aes(x=Time_days,y=Akkermansia + 10^-6, col=cluster %>% as.character)) + guides(col=FALSE) +
  geom_vline(xintercept = 32) + geom_hline(yintercept = 0.009129289) + ylab("Akkermansia")



```

## akkermansia over time clustering


```{r}

mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  filter(DSS_group=="High dose") %>%
  select(Akkermansia,Time_days,Rat_ID) %>%
  reshape2::dcast(Rat_ID~Time_days, value.var = "Akkermansia", fill=0) %>%
  tibble::column_to_rownames("Rat_ID") %>%
  kmeans(.,3) %>%
  .$cluster %>%
  as.matrix %>%
  as.data.frame %>%
  rename(cluster=V1) -> rat_cluster

hysteresis = read.csv2("hysteresis.csv") %>% 
  mutate(hysteresis_grp = fct_recode(hysteresis_grp, `resistant\nprofile` = "high",
                                                      `resilient\nprofile` = "intermediate",                          
                                                      `no resilient\nprofile` = "low"))

mapping_file %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(Rat_ID,samplenr,Time,DSS,Study_group) %>%
  filter(Study_group != "G1") %>%
  mutate(Time_days = Time %>% gsub("Tm","-", .) %>% gsub("T","",.) %>% as.character %>% as.numeric) %>% filter(Time_days >= -1) %>% 
  filter(Time != "T47") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  merge(.,t(genus[, select_DSS_samples %>% as.character]), by.x="samplenr", by.y="row.names") %>%
  select(contains("Akkermansia"),DSS,Time_days, Rat_ID) %>%
  mutate(DSS_group = ifelse(DSS>=1,"High dose","No or Low dose")) %>%
  #filter(DSS ==0) %>%
  rename_at(1, ~"Akkermansia") %>%
  filter(DSS_group=="High dose") %>%
  select(Akkermansia,Time_days,Rat_ID) %>%
  merge(.,rat_cluster,by.x="Rat_ID",by.y="row.names") %>%
  merge(.,hysteresis,by="Rat_ID") %>%
  ggplot() + 
  geom_line(aes(x=Time_days,y=Akkermansia + 10^-6, group=Rat_ID), linetype=3) + 
  geom_point(aes(x=Time_days,y=Akkermansia + 10^-6,  col=hysteresis_grp %>% as.character)) +
  scale_y_log10() + facet_wrap(~hysteresis_grp) + 
  geom_smooth(aes(x=Time_days,y=Akkermansia + 10^-6, col=hysteresis_grp %>% as.character)) + guides(col=FALSE) +
  geom_vline(xintercept = 32) + geom_hline(yintercept = 0.009129289) + ylab("Akkermansia")




```





# figure stability potential

```{r, fig.height=6, fig.width=10}


source("PlotPotential2.r")

test = merge(genus_pco$li, 
             mapping_file %>% 
               merge(., rat_metadata, by.x="Rat_ID", by.y="Animal_id_2"), by.x="row.names", by.y="X.SampleID") %>% 
  #merge(.,data.frame(observed_otu, microbiota_type = as.character(microbiota_type)), by.x="Row.names", by.y="row.names") %>% 
  filter(Time %in% c("T68") ) %>% 
  filter(Study_group != "G1") %>%
  mutate(DSS = Study_group %>% forcats::fct_recode(`0` = "G2",`0.25` = "G3",`0.5` = "G4",`1` = "G5",`2` = "G6",`3` = "G7") %>% as.character %>% as.numeric ) %>%
  filter(DSS != 0) %>%
  select(DSS,A1,A2)


res2 <- movpotential_ews(test$A1, test$DSS  %>% as.character %>% as.numeric)


PlotPotential2(res2$res, title = '', 
               ylab.text = 'Microbial State\n(JSD derived metric)', xlab.text = 'Chronical Perturbation strength (DSS %)', 
           cutoff = 0.5, plot.contours = TRUE, binwidth = 0.1) + theme_classic() 



PlotPotential2_df =function(res, title = "", xlab.text, ylab.text, cutoff = 0.5, 
          plot.contours = TRUE, binwidth = 0.2, bins = NULL) {
  
   cut.potential <- max(apply(res$pots, 1, min)) + cutoff * 
    abs(max(apply(res$pots, 1, min)))
  pots <- res$pots
  pots[pots > cut.potential] <- cut.potential
  intp <- tgp::interp.loess(as.vector(res$pars), as.vector(res$xis), 
                            as.vector(pots), gridlen = 2 * dim(pots))
  xy <- expand.grid(intp$x, intp$y)
  z <- as.vector(intp$z)
  z[is.na(z)] <- max(na.omit(z))
  potential <- NULL
  df <- data.frame(list(bg.var = xy[, 1], phylotype = xy[, 
                                                         2], potential = z))
  
  
  return(df)
  
          }

PlotPotential2_plot = function(df, binwidth = 0.2, bins = NULL, plot.contours = TRUE, xlab.text, ylab.text, title = ""){
  
  bg.var <- NULL
  phylotype <- NULL
  p <- ggplot2::ggplot(df, aes(bg.var, phylotype, z = potential)) + 
    geom_tile(aes(fill = potential)) + scale_fill_gradientn(colours = topo.colors(10))
  if (plot.contours) {
    if (!is.null(bins)) {
      warning("bins argument is overriding the binwidth argument!")
      p <- p + stat_contour(bins = bins)
    }
    else {
      p <- p + stat_contour(binwidth = binwidth)
    }
  }
  p <- p + xlab(xlab.text) + ylab(ylab.text) + labs(title = title)
  p
  
}

PlotPotential2_df(res2$res) %>% 
  PlotPotential2_plot(title = '', ylab.text = 'Microbial State\n(JSD derived metric)', xlab.text = 'Chronical Perturbation strength (DSS %)')


PlotPotential2_df(res2$res)






l1 = wireframe(
  potential~bg.var+phylotype, 
  data = PlotPotential2_df(res2$res), 
  drape = TRUE, shade=FALSE, aspect = c(0.8, 0.4), 
  light.source = c(10,0,10), 
  col.regions = topo.colors(100), 
  screen = list(z = 10, x = -40, y=10), colorkey=FALSE, 
  xlab="Perturbation\nstrength\n(DSS)", ylab="Ecological\nstate\n(JSD  metrics)", zlab="Potential",
  par.box = list(col="grey"),col="grey",lwd=0.1, 
  scales=list(arrows=F,col=1),
  par.settings = 
    list(axis.line = list(col = "transparent"),
         
         layout.heights =
        list(top.padding = 0,
 	    main.key.padding = 0,
 	    key.axis.padding = 0,
 	    axis.xlab.padding = 0,
 	    xlab.key.padding = 0,
 	    key.sub.padding = 0,
 	    bottom.padding = 0),
 	    
 	    
 	      layout.widths =
        list(left.padding = 0,
 	    key.ylab.padding = 0,
 	    ylab.axis.padding = 0,
 	    axis.key.padding = 0,
 	    right.padding = 0))
 	    
         )


print(l1)


g1 <- grid::grid.grabExpr(print(l1)) 

plot_grid(g1,pA_akk, nrow=2, labels = "AUTO") 


ggsave("alternative_state.pdf",h=15,w=10)


PlotPotential2_df(res2$res) %>% dim

```



```{r, eval=FALSE, include=FALSE}

pdf(file="plot_potential_animate.pdf", width=4, height=4)
  for (i in seq(0, 360 ,3)){
    plot_potential=
    wireframe(
  potential~bg.var+phylotype, 
  data = PlotPotential2_df(res2$res), 
  drape = TRUE, shade=FALSE, aspect = c(0.8, 0.4), 
  light.source = c(10,0,10), 
  col.regions = topo.colors(100), 
  screen = list(z = i, x = -40, y=10), colorkey=FALSE, 
  xlab="Perturbation\nstrength\n(DSS)", ylab="Ecological\nstate\n(JSD  metrics)", zlab="Potential",
  par.box = list(col="grey"),col="grey",lwd=0.1, 
  scales=list(arrows=F,col=1),
  par.settings = 
    list(axis.line = list(col = "transparent"),
         
         layout.heights =
        list(top.padding = 0,
 	    main.key.padding = 0,
 	    key.axis.padding = 0,
 	    axis.xlab.padding = 0,
 	    xlab.key.padding = 0,
 	    key.sub.padding = 0,
 	    bottom.padding = 0),
 	    
 	    
 	      layout.widths =
        list(left.padding = 0,
 	    key.ylab.padding = 0,
 	    ylab.axis.padding = 0,
 	    axis.key.padding = 0,
 	    right.padding = 0))
 	    
         )
    
    print(plot_potential)
  }
dev.off()

# convert pdf to gif using ImageMagick
#system("convert -delay 40 *.pdf example_3_reduced.gif")





```


## hysteresis and resilience



import histological data
```{r, fig.height=7, fig.width=10}


hysteresis = read.csv2("hysteresis.csv") %>% 
  mutate(hysteresis_grp = fct_recode(hysteresis_grp, `resistant\nprofile` = "high",
                                                      `resilient\nprofile` = "intermediate",                          
                                                      `no resilient\nprofile` = "low"))


histo <- readxl::read_excel("Richness-2017-03-23-Data Transfer Matrix-SDB.xlsx", 
    sheet = "Histopathological reading", 
    skip = 2)

colnames(histo)[3] = "Rat_ID"

histo %>% .[,c(3,28,39)] %>% 
  #rename_at(1,~"Rat_ID") %>%
  rename(score_histo=`Total (max=37)`) %>%
  filter(Region=="distal") %>%
  mutate(Rat_ID=paste0("R",Rat_ID)) -> histo 




```


build data.frame with akkermansia and time point
```{r}



mapping_file %>% 
  merge(
    ., 
    rat_metadata, 
    by.x="Rat_ID", 
    by.y="Animal_id_2"
    ) %>% 
  merge(.,hysteresis, by = "Rat_ID") %>%
  merge(., 
        genus[row.names(genus) %>% grep("Akkermansia",.),] %>% as.data.frame(row.names="genus") %>% t , 
        #genus[row.names(genus) %>% grep("Ruminococcus",.),] %>% as.data.frame(row.names="genus") %>% t ,
        by.x="X.SampleID", 
        by.y="row.names") %>% 
  merge(.,genus_pco$li, by.y="row.names", by.x="X.SampleID") %>%
select(X.SampleID, A1,A2,A3, genus, hysteresis_grp, Time,Study_group,Rat_ID) %>%
  mutate(Time=fct_relevel(Time, c("Tm49","Tm41","Tm34","Tm20","Tm13","Tm7","Tm1","T3","T13","T17","T27","T31","T40","T47","T63","T68"))) %>%
  mutate(DSSpulse = fct_recode(Time, 
                            Water = "Tm49",  
                            Water = "Tm41", 
                            Water = "Tm34", 
                            Water = "Tm20", 
                            Water = "Tm13", 
                            Water = "Tm7", 
                            DSS = "Tm1", 
                            Water   = "T3",
                            DSS = "T13",
                            Water   = "T17",
                            DSS = "T27",
                            Water   = "T31",
                            Water = "T40",
                            Water = "T47",
                            Water = "T63",
                            Water = "T68"
                            
                            )) %>%
  
mutate(diet = fct_recode(Time, 
                            MZ  = "Tm49",  
                            MZ  = "Tm41", 
                            MZ  = "Tm34", 
                            AIN = "Tm20", 
                            AIN = "Tm13", 
                            AIN = "Tm7", 
                            AIN = "Tm1", 
                            AIN = "T3",
                            AIN = "T13",
                            AIN = "T17",
                            AIN = "T27",
                            AIN = "T31",
                            AIN = "T40",
                            AIN = "T47",
                            AIN = "T63",
                            AIN = "T68"
                            
                            )) %>%
  
  mutate(perturbation = fct_recode(Time,
                            `0` = "Tm1", 
                            `3` = "T3",
                            `0` = "T13",
                            `3` = "T17",
                            `0` = "T27",
                            `3` = "T31",
                            `0` = "T40",
                            `-3` = "T47",
                            `-6` = "T63",
                            `-8` = "T68"
                            
                            ))  %>%
  
  mutate(press_cumul_day = fct_recode(Time,
                            `0` = "Tm1", 
                            `3` = "T3",
                            `13` = "T13",
                            `17` = "T17",
                            `27` = "T27",
                            `31` = "T31",
                            `22` = "T40",
                            `15` = "T47",
                            `0` = "T63",
                            `-6` = "T68"
                            
                            ))  %>%
  
mutate(DSSpress = fct_recode(Time, 
                            Water = "Tm49",  
                            Water = "Tm41", 
                            Water = "Tm34", 
                            Water = "Tm20", 
                            Water = "Tm13", 
                            Water = "Tm7", 
                            DSS = "Tm1", 
                            DSS   = "T3",
                            DSS = "T13",
                            DSS   = "T17",
                            DSS = "T27",
                            DSS   = "T31",
                            Water = "T40",
                            Water = "T47",
                            Water = "T63",
                            Water = "T68"
                            
                            )) %>%
  
  mutate(Time=Time %>% gsub("T","",.) %>% gsub("m","-",.) %>% as.character %>% as.numeric) %>%
  filter(Time > -2) %>%
  mutate(perturbation = perturbation %>% as.character %>% as.numeric) %>%
  filter(Study_group != "G1") %>%
  arrange(Time) -> akk_df


save(akk_df, file="akk_df.rda")



```




```{r, fig.height=4, fig.width=10}

resilience_plot =

akk_df %>%
  filter(Study_group %in% c("G5","G6","G7")) %>%
  ggplot() + 
  #geom_point(aes(x=Time,y=genus+10^-6)) + 
  geom_smooth(aes(x=Time,y=genus+10^-6, group=hysteresis_grp)) + 
  facet_wrap(~hysteresis_grp) + 
  scale_y_log10() + geom_hline(yintercept = 10^-1.98, linetype =2) +
  geom_vline(xintercept = 31, linetype =2) + ylab("Akkermansia") + xlab("Time (days)") +
  annotate("text",x=15,y=0.3,label="DSS pulse\n(x3)", size=3) +
  annotate("text",x=45,y=0.3,label="Washout\n", size=3) +
  annotate("text",x=5,y=0.015,label="Tipping point", size=3) 

resilience_plot

```


```{r, fig.height=4, fig.width=3}



histo_DSS_hysteresis_barplot =
akk_df %>%
  #filter(Study_group != "G2") %>%
  select(hysteresis_grp,Study_group, Rat_ID) %>% 
  unique %>%
  merge(.,histo, by="Rat_ID") %>%
  mutate(DSS_group = ifelse(Study_group == "G5" | Study_group == "G6" | Study_group == "G7", "high", "low" )) %>%
  filter(DSS_group == "high") %>%
  mutate(score_histo = as.numeric(as.character(score_histo))) %>%
  group_by(hysteresis_grp) %>%
  summarise(n = n(), m = mean(score_histo), sem = sd(score_histo)/sqrt(n())) %>%
  ungroup() %>%
  ggplot(aes(x=hysteresis_grp, y=m)) +
  geom_bar(fill="white", stat="identity", position="dodge", color="black") + 
  geom_errorbar(aes(ymin=m-sem,ymax=m+sem),width=.2, position=position_dodge(.9)) +
  #scale_fill_brewer(type="qual") +
  ylab("Histological score") +
  xlab("") +
  theme(legend.position = c(0.1, 0.9))


histo_DSS_hysteresis_barplot


```



```{r, fig.height=10, fig.width=10}

plot_grid(resilience_plot,plot_grid(g1,histo_DSS_hysteresis_barplot,nrow=1,rel_widths = c(4,1.5), labels=c("B","C")),nrow=2,rel_heights = c(1,1.6), labels = c("A",NA))

ggsave("Figure_Akk_hysteresis_host.pdf")

```



## Akkermansia network analysis



### microbial interaction with akkermansia

```{r, fig.width=15, message=FALSE, warning=FALSE}


mapping_file %>% 
  merge(
    ., 
    rat_metadata, 
    by.x="Rat_ID", 
    by.y="Animal_id_2"
    ) %>% 
  merge(.,hysteresis, by = "Rat_ID", all.x=TRUE) %>%
  merge(., 
        genus %>% noise.removal(percent=1) %>% t, 
        by.x="X.SampleID", 
        by.y="row.names") %>%
  mutate(Time=fct_relevel(Time, c("Tm49","Tm41","Tm34","Tm20","Tm13","Tm7","Tm1","T3","T13","T17","T27","T31","T40","T47","T63","T68"))) %>% 
  mutate(Time=Time %>% gsub("T","",.) %>% gsub("m","-",.) %>% as.character %>% as.numeric) %>%
  
  select(-X.SampleID, -BarcodeSequence, -LinkerPrimerSequence, -samplenr, -INRA_DNA_ID, -DNAextractionOperator, -Time_DSS, -Description, -Animal_id, -Weight, -Bacteroides, -Filiation, -Cage_provider, -Animal.Identification ) %>%
  filter(Diet != "AIN93G", Time > -10) %>%
  filter(DSS == "1" | DSS == "2" | DSS == "3") %>%
  arrange(Study_group, Rat_ID,Time) %>%
  melt(id.vars=c("Rat_ID","Time","DSS","Diet","Study_group","hysteresis_grp","D_0__Bacteria;D_1__Verrucomicrobia;D_2__Verrucomicrobiae;D_3__Verrucomicrobiales;D_4__Verrucomicrobiaceae;D_5__Akkermansia")) %>%
  rename(Akkermansia = `D_0__Bacteria;D_1__Verrucomicrobia;D_2__Verrucomicrobiae;D_3__Verrucomicrobiales;D_4__Verrucomicrobiaceae;D_5__Akkermansia`) %>%
  group_by(Rat_ID,variable,hysteresis_grp) %>%
  summarise(akk_cor=cor(Akkermansia,value, method="spearman")) %>%
  mutate(akk_cor = ifelse(is.na(akk_cor),0,akk_cor)) %>%
  ungroup() -> akk_cor



akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) ) %>%
  mutate(direction = ifelse(median>0,"pos","neg")) %>%
  filter((direction == "pos" & q25 > 0)|(direction == "neg" & q75 < 0)) %>%
  filter(abs(median) > 0.4) %>%
  merge(.,akk_cor, by=c("hysteresis_grp","variable"), all.x=TRUE) %>%
  tidyr::separate(variable,into=c("Domain","Phylum","Class","Order","Family","Genus"),sep=";") %>%
  filter(Genus != "Other", Family != "D_5__uncultured") %>%
  ungroup() %>%
  ggplot() + geom_boxplot(aes(x=paste(Family,Genus),y=akk_cor, fill=median)) + facet_wrap(~hysteresis_grp) +
  coord_flip() +
  scale_fill_continuous(low = "#FF0000", high="#00FF00") +
  ylab("Genus correlation over time with\nAkkermansia rel. abundance") + xlab("")




akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) )




```



```{r, fig.height=6, fig.width=20}

library(ggraph)
library(igraph)
library(tidygraph)

akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) ) %>%
  filter(abs(median) > 0.4) %>%
  merge(.,akk_cor, by=c("hysteresis_grp","variable"), all.x=TRUE) %>%
  tidyr::separate(variable,into=c("Domain","Phylum","Class","Order","Family","Genus"),sep=";") %>%
  filter(Genus != "Other", Family != "D_5__uncultured") %>%
  ungroup()

my_lab <- function(labels) {
  list(hysteresis_grp = c("resistant\nprofile","resilient\nprofile","no resilient\nprofile"))
}

akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) ) %>%
  mutate(direction = ifelse(median>0,"pos","neg")) %>%
  filter((direction == "pos" & q25 > 0)|(direction == "neg" & q75 < 0)) %>%
  #filter(abs(median) > 0.25) %>%
   tidyr::separate(variable,into=c("Domain","Phylum","Class","Order","Family","Genus"),sep=";") %>%
  filter(Genus != "Other", Family != "D_5__uncultured") %>%
  mutate(Akkermansia="Akkermansia") %>%
  select(Akkermansia,Genus,hysteresis_grp,median) %>%
  ungroup() %>%
  #mutate(hysteresis_grp = hysteresis_grp %>% factor(levels=c("resistant\nprofile","resilient\nprofile","no resilient\nprofile"))) %>%
  mutate(Genus = Genus %>% gsub("D_5__","",.)) %>%
  igraph::graph_from_data_frame() %>%
  ggraph(., layout = 'kk') + 
    geom_edge_link(aes(colour = median), width=2, alpha=0.7) + 
  facet_edges(~hysteresis_grp , nrow=1, labeller = my_lab) +   
  geom_node_text(aes(label = name)) +
  scale_edge_colour_gradient2(low = "red", mid="yellow", high="green") +
  theme_dark()
  

#%>%
#  ggraph() + geom_edge


akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) ) %>%
  mutate(direction = ifelse(median>0,"pos","neg")) %>%
  filter((direction == "pos" & q25 > 0)|(direction == "neg" & q75 < 0)) %>%
  filter(abs(median) > 0.4) %>%
   tidyr::separate(variable,into=c("Domain","Phylum","Class","Order","Family","Genus"),sep=";") %>%
  mutate(Genus = ifelse(Family == "D_4__Prevotellaceae" & Genus == "D_5__uncultured","Prevotellaceae",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__Ruminococcaceae" & Genus == "D_5__uncultured","Ruminococcaceae",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__Erysipelotrichaceae" & Genus == "D_5__Incertae Sedis","Erysipelotrichaceae",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__S24-7" & Genus == "D_5__uncultured","S24-7a",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__Family XIII" & Genus == "D_5__uncultured","Clostridiales XIII",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__S24-7" & Genus == "Other","S24-7b",Genus)) %>%
   mutate(Genus = ifelse(Family == "D_4__S24-7" & Genus == "Ambiguous_taxa","S24-7c",Genus)) %>%
  filter(Genus != "Other", Family != "D_4__uncultured") %>%
  mutate(Akkermansia="Akkermansia") %>%
  select(Akkermansia,Genus,hysteresis_grp,median) %>%
  ungroup() %>%
  #mutate(hysteresis_grp = hysteresis_grp %>% factor(levels=c("resistant\nprofile","resilient\nprofile","no resilient\nprofile"))) %>%
  mutate(Genus = Genus %>% gsub("D_5__","",.)) %>%
  igraph::graph_from_data_frame() -> akk_net

edge.attributes(akk_net)$hysteresis_grp = edge.attributes(akk_net)$hysteresis_grp %>% factor(levels=c("resistant\nprofile","resilient\nprofile","no resilient\nprofile"))


akk_net %>%
ggraph(., layout = 'kk') + 
    geom_edge_link(aes(colour = median), width=2, alpha=0.5) + 
  facet_edges(~hysteresis_grp , nrow=1, labeller = my_lab) +   
  geom_node_text(aes(label = name)) +
  scale_edge_colour_gradient2(low = "red", mid="yellow", high="green") +
  theme_dark() + theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
                       axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
                       strip.text.x = element_text(size = 20)) + 
  xlab("") + ylab("") -> akk_solo_plot

akk_solo_plot

ggsave("akk_solo_plot.pdf")

```

focus on Subdoligranulum


```{r}



genera_select_melt %>%
  filter(variable == "D_0__Bacteria;D_1__Firmicutes;D_2__Clostridia;D_3__Clostridiales;D_4__Ruminococcaceae;D_5__Subdoligranulum") %>%
  ggplot() + 
  geom_smooth(aes(x=Time,y=value+10^-6, group=hysteresis_grp)) + 
  facet_wrap(~hysteresis_grp) + 
  scale_y_log10() + #geom_hline(yintercept = 10^-1.98, linetype =2) +
  geom_vline(xintercept = 31, linetype =2) + ylab("Subdoligranulum") + xlab("Time (days)") +
  annotate("text",x=15,y=0.001,label="DSS pulse\n(x3)", size=3) +
  annotate("text",x=45,y=0.001,label="Washout\n", size=3) #+
  #annotate("text",x=5,y=0.015,label="Tipping point", size=3) 
  

select_genera

```




### global network around Akkermansia

```{r, fig.height=3, fig.width=10}

select_genera = 

akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) ) %>%
  mutate(direction = ifelse(median>0,"pos","neg")) %>%
  filter((direction == "pos" & q25 > 0)|(direction == "neg" & q75 < 0)) %>% 
  filter(abs(median) > 0.4) %>%
  mutate(variable = variable %>% as.character) %>% .$variable %>% unique


select_genera =c("D_0__Bacteria;D_1__Verrucomicrobia;D_2__Verrucomicrobiae;D_3__Verrucomicrobiales;D_4__Verrucomicrobiaceae;D_5__Akkermansia",select_genera )


mapping_file %>% 
  merge(
    ., 
    rat_metadata, 
    by.x="Rat_ID", 
    by.y="Animal_id_2"
    ) %>% 
  merge(.,hysteresis, by = "Rat_ID", all.x=TRUE) %>%
  merge(., 
        genus %>% .[select_genera,] %>% t, 
        by.x="X.SampleID", 
        by.y="row.names") %>%
  mutate(Time=fct_relevel(Time, c("Tm49","Tm41","Tm34","Tm20","Tm13","Tm7","Tm1","T3","T13","T17","T27","T31","T40","T47","T63","T68"))) %>% 
  mutate(Time=Time %>% gsub("T","",.) %>% gsub("m","-",.) %>% as.character %>% as.numeric) %>%
  
  select(-X.SampleID, -BarcodeSequence, -LinkerPrimerSequence, -samplenr, -INRA_DNA_ID, -DNAextractionOperator, -Time_DSS, -Description, -Animal_id, -Weight, -Bacteroides, -Filiation, -Cage_provider, -Animal.Identification ) %>%
  filter(Diet != "AIN93G", Time > -10) %>%
  filter(DSS == "1" | DSS == "2" | DSS == "3") %>%
  select(-Diet, -Study_group) %>%
  melt(id.vars=c("Rat_ID","Time","DSS","hysteresis_grp")) -> genera_select_melt


genera_select_melt %>%
  merge(.,genera_select_melt,by=c("Rat_ID","Time","DSS","hysteresis_grp")) %>%
  filter(!(variable.x==variable.y)) %>%
  group_by(Rat_ID,DSS,hysteresis_grp,variable.x,variable.y) %>%
  summarise(rho = cor(value.x,value.y,method="spearman") ) -> network_df_cor

network_df_cor %>%
  na.omit() %>%
  ungroup() %>%
  group_by(hysteresis_grp,variable.x,variable.y) %>%
  summarise(n=n(), median=median(rho), q25=quantile(rho,0.25), q75 = quantile(rho,0.75) ) %>%
  mutate(direction = ifelse(median>0,"pos","neg")) %>%
  filter((direction == "pos" & q25 > 0)|(direction == "neg" & q75 < 0)) %>% 
  filter(abs(median) > 0.4) %>%
  tidyr::separate(variable.x,into=paste0(c("Domain","Phylum","Class","Order","Family","Genus"),".x"),sep=";") %>%
  tidyr::separate(variable.y,into=paste0(c("Domain","Phylum","Class","Order","Family","Genus"),".y"),sep=";") %>%
  mutate(Genus.x = ifelse(Family.x == "D_4__Prevotellaceae" & Genus.x == "D_5__uncultured","Prevotellaceae",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__Prevotellaceae" & Genus.y == "D_5__uncultured","Prevotellaceae",Genus.y)) %>%
  mutate(Genus.x = ifelse(Family.x == "D_4__Ruminococcaceae" & Genus.x == "D_5__uncultured","Ruminococcaceae",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__Ruminococcaceae" & Genus.y == "D_5__uncultured","Ruminococcaceae",Genus.y)) %>%
  mutate(Genus.x = ifelse(Family.x == "D_4__Erysipelotrichaceae" & Genus.x == "D_5__Incertae Sedis","Erysipelotrichaceae",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__Erysipelotrichaceae" & Genus.y == "D_5__Incertae Sedis","Erysipelotrichaceae",Genus.y)) %>%
  mutate(Genus.x = ifelse(Family.x == "D_4__S24-7" & Genus.x == "D_5__uncultured","S24-7a",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__S24-7" & Genus.y == "D_5__uncultured","S24-7a",Genus.y)) %>%
  mutate(Genus.x = ifelse(Family.x == "D_4__Family XIII" & Genus.x == "D_5__uncultured","Clostridiales XIII",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__Family XIII" & Genus.y == "D_5__uncultured","Clostridiales XIII",Genus.y)) %>%
  mutate(Genus.x = ifelse(Family.x == "D_4__S24-7" & Genus.x == "Other","S24-7b",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__S24-7" & Genus.y == "Other","S24-7b",Genus.y)) %>%
   mutate(Genus.x = ifelse(Family.x == "D_4__S24-7" & Genus.x == "Ambiguous_taxa","S24-7c",Genus.x),
         Genus.y = ifelse(Family.y == "D_4__S24-7" & Genus.y == "Ambiguous_taxa","S24-7c",Genus.y)) %>%
  filter(Genus.x != "Other", Family.x != "D_4__uncultured",Genus.y != "Other", Family.y != "D_4__uncultured") %>%
  select(Genus.x,Genus.y,hysteresis_grp,median) %>%
  ungroup() %>%
  #mutate(hysteresis_grp = hysteresis_grp %>% factor(levels=c("resistant\nprofile","resilient\nprofile","no resilient\nprofile"))) %>%
  mutate(Genus.x = Genus.x %>% gsub("D_5__","",.), Genus.y = Genus.y %>% gsub("D_5__","",.)) %>%
  igraph::graph_from_data_frame() -> global_akk_net

edge.attributes(global_akk_net)$hysteresis_grp = edge.attributes(global_akk_net)$hysteresis_grp %>% factor(levels=c("resistant\nprofile","resilient\nprofile","no resilient\nprofile"))


global_akk_net_plot =

global_akk_net %>%
ggraph(., layout = 'kk') + 
    geom_edge_link(aes(colour = median),alpha=0.5, width=0.5) + 
  facet_edges(~hysteresis_grp , nrow=1) +   
  geom_node_text(aes(label = name),size=3) +
  scale_edge_colour_gradient2(low = "red", mid="yellow", high="green") +
  theme_dark() + theme(axis.text.x = element_blank(), axis.text.y = element_blank(),
                       axis.ticks.x=element_blank(), axis.ticks.y=element_blank(),
                       strip.text.x = element_text(size = 10),
                       panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank()) + 
  xlab("") + ylab("") 

global_akk_net_plot
ggsave("global_akk_net.pdf")

```



focus on Subdoligranulum


```{r}



genera_select_melt %>%
  filter(variable == "D_0__Bacteria;D_1__Firmicutes;D_2__Clostridia;D_3__Clostridiales;D_4__Ruminococcaceae;D_5__Subdoligranulum") %>%
  ggplot() + 
  geom_smooth(aes(x=Time,y=value+10^-6, group=hysteresis_grp)) + 
  facet_wrap(~hysteresis_grp) + 
  scale_y_log10() + #geom_hline(yintercept = 10^-1.98, linetype =2) +
  geom_vline(xintercept = 31, linetype =2) + ylab("Subdoligranulum") + xlab("Time (days)") +
  annotate("text",x=15,y=0.001,label="DSS pulse\n(x3)", size=3) +
  annotate("text",x=45,y=0.001,label="Washout\n", size=3) #+
  #annotate("text",x=5,y=0.015,label="Tipping point", size=3) 
  



```





```{r, fig.height=30, fig.width=7}

genera_select_melt %>%
  ggplot() + 
  geom_smooth(aes(x=Time,y=value+10^-6, group=hysteresis_grp)) + 
  facet_grid(variable %>% gsub(";","\n",.)~hysteresis_grp, scale="free_y") + 
  theme(strip.text.y = element_text(size = 7)) +
  scale_y_log10() #+ #geom_hline(yintercept = 10^-1.98, linetype =2) +
  #geom_vline(xintercept = 31, linetype =2) + #ylab("Alloprevotella") + xlab("Time (days)") +
  #annotate("text",x=15,y=0.001,label="DSS pulse\n(x3)", size=3) +
  #annotate("text",x=45,y=0.001,label="Washout\n", size=3) #+
  #annotate("text",x=5,y=0.015,label="Tipping point", size=3) 

  
  


```


```{r, fig.height=8, fig.width=10}


select_genera = 

akk_cor%>%
  group_by(hysteresis_grp,variable) %>%
  summarise(n=n(), median=median(akk_cor), q25=quantile(akk_cor,0.25), q75 = quantile(akk_cor,0.75) ) %>%
  mutate(direction = ifelse(median>0,"pos","neg")) %>%
  filter((direction == "pos" & q25 > 0)|(direction == "neg" & q75 < 0)) %>% 
  filter(abs(median) > 0.4) %>%
  mutate(variable = variable %>% as.character) %>% .$variable %>% unique


select_genera =c("D_0__Bacteria;D_1__Verrucomicrobia;D_2__Verrucomicrobiae;D_3__Verrucomicrobiales;D_4__Verrucomicrobiaceae;D_5__Akkermansia",select_genera )


mapping_file %>% 
  merge(
    ., 
    rat_metadata, 
    by.x="Rat_ID", 
    by.y="Animal_id_2"
    ) %>% 
  merge(.,hysteresis, by = "Rat_ID", all.x=TRUE) %>%
  merge(., 
        genus %>% .[select_genera,] %>% t, 
        by.x="X.SampleID", 
        by.y="row.names") %>%
  mutate(Time=fct_relevel(Time, c("Tm49","Tm41","Tm34","Tm20","Tm13","Tm7","Tm1","T3","T13","T17","T27","T31","T40","T47","T63","T68"))) %>% 
  mutate(Time=Time %>% gsub("T","",.) %>% gsub("m","-",.) %>% as.character %>% as.numeric) %>%
  
  select(-X.SampleID, -BarcodeSequence, -LinkerPrimerSequence, -samplenr, -INRA_DNA_ID, -DNAextractionOperator, -Time_DSS, -Description, -Animal_id, -Weight, -Bacteroides, -Filiation, -Cage_provider, -Animal.Identification ) %>%
  filter(Diet != "AIN93G", Time > -10) %>%
  filter(DSS == "1" | DSS == "2" | DSS == "3") %>%
  select(-Diet, -Study_group) %>%
  melt(id.vars=c("Rat_ID","Time","DSS","hysteresis_grp")) -> genera_select_melt


# 
#   genera_select_melt %>%
#   group_by(variable,hysteresis_grp, Time) %>%
#   summarise(value = (value+10^-6) %>% log10() %>% mean() ) %>%
#   ungroup() %>%
#   group_by(variable,Time) %>%
#   mutate(value = value %>% scale()) %>%
#   ungroup() %>%
#   reshape2::dcast(variable~hysteresis_grp+Time) %>%
#   tibble::column_to_rownames("variable") %>%
#   kmeans(.,3) %>%
#   .$cluster %>%
#   as.matrix %>%
#   as.data.frame %>%
#   rename(cluster=V1) -> genus_cluster


  
  genera_select_melt %>%
  group_by(variable,hysteresis_grp, Time) %>%
  summarise(value = (value+10^-6) %>% log10() %>% mean() ) %>%
  ungroup() %>%
  group_by(variable,Time) %>%
  mutate(value = value %>% scale()) %>%
  ungroup() %>%
  reshape2::dcast(variable~hysteresis_grp+Time) %>%
  tibble::column_to_rownames("variable") %>% 
    fpc::pamk( metric="euclidean") %>%
    .$pamobject %>%
    .$clustering %>%
    as.matrix() %>%
    as.data.frame() %>%
    rename(cluster = V1) -> genus_cluster 

  
  
  
 

  
pp = genera_select_melt %>%
  merge(.,genus_cluster, by.x="variable", by.y="row.names") %>%
  group_by(cluster,variable,hysteresis_grp, Time) %>%
  summarise(value = (value+10^-6) %>% log10() %>% mean() ) %>%
  ungroup() %>%
  group_by(variable,Time) %>%
  mutate(value = value  %>% scale() ) %>%
  ggplot() + geom_boxplot(aes(x=Time%>%as.character,y=value)) +
  facet_grid(cluster~hysteresis_grp, scale="free_y") + 
  theme(strip.text.y = element_text(size = 7)) 
  
  
 #g=gridExtra::tableGrob(genus_cluster), theme=gridExtra::ttheme_minimal())

 
mytheme <- gridExtra::ttheme_default(
    core = list(fg_params=list(cex = 0.5)),
    colhead = list(fg_params=list(cex = 0.5)),
    rowhead = list(fg_params=list(cex = 0.5)))

 genus_cluster %>% 
tibble::rownames_to_column("genus") %>% 
 tidyr::separate(genus,into=c("Domain","Phylum","Class","Order","Family","Genus"),sep=";") %>%
   mutate(Genus = ifelse(Family == "D_4__Prevotellaceae" & Genus == "D_5__uncultured","Prevotellaceae",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__Ruminococcaceae" & Genus == "D_5__uncultured","Ruminococcaceae",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__Erysipelotrichaceae" & Genus == "D_5__Incertae Sedis","Erysipelotrichaceae",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__S24-7" & Genus == "D_5__uncultured","S24-7a",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__Family XIII" & Genus == "D_5__uncultured","Clostridiales XIII",Genus)) %>%
  mutate(Genus = ifelse(Family == "D_4__S24-7" & Genus == "Other","S24-7b",Genus)) %>%
   mutate(Genus = ifelse(Family == "D_4__S24-7" & Genus == "Ambiguous_taxa","S24-7c",Genus)) %>%
   select(Genus,cluster) %>%
   mutate(Genus = Genus %>% gsub("D_5__","",.)) %>%
   arrange(cluster) %>%
   gridExtra::tableGrob(theme = mytheme) -> genus_clusters_table

genus_dynamic_table = genera_select_melt %>%
  merge(.,genus_cluster, by.x="variable", by.y="row.names") %>%
  group_by(variable,Time) %>%
  mutate(value = (value+10^-6)  %>% log10() %>% scale() ) %>%
  ggplot() +
  geom_smooth(aes(x=Time,y=value+10^-6, group=hysteresis_grp), method="loess") + 
  facet_grid(cluster~hysteresis_grp, scale="free_y") + 
  theme(strip.text.y = element_text(size = 7)) + ylab("Rel. abundance\nscaled") +
  geom_vline(xintercept = 31, linetype = 2)

plot_grid(genus_dynamic_table,genus_clusters_table, ncol=2, rel_widths = c(2,1))

```

### global figure

```{r, fig.height=12, fig.width=12}

plot_grid(

plot_grid(
resilience_plot,
histo_DSS_hysteresis_barplot, nrow=1, rel_widths = c(3,1)),
plot_grid(NULL,global_akk_net_plot,NULL, rel_widths = c(0.3,4,1), nrow=1),
plot_grid(genus_dynamic_table,
genus_clusters_table, nrow=1, rel_widths = c(3,1)), nrow=3, rel_heights = c(1.5,1.5,3))

ggsave("resilience_network_plot.pdf")


```




## import gro alpha data

```{r, eval=FALSE, include=FALSE}
GRO_alpha = readxl::read_excel("Multiplex_Richness_SDB_200318.xlsx", sheet = "Multiplex T41 T47", skip=2) %>%
  select(3,`Blood harvest Timepoint`,33) %>%
  filter(`Blood harvest Timepoint` == "T47") %>%
  na.omit

colnames(GRO_alpha) = c("Rat_ID","Time","GRO_Alpha")
  

GRO_alpha %>%
  mutate(Rat_ID = paste0("R",Rat_ID)) %>%
  mutate(GRO_Alpha = gsub("<=0","0",GRO_Alpha) %>% as.numeric) -> GRO_alpha
  
  
  

```


```{r, eval=FALSE, include=FALSE}

GRO_alpha %>%
  merge(.,rat_metadata, by.x="Rat_ID", by.y="Animal_id_2") %>%
  merge(hysteresis, by="Rat_ID") %>%
  select(Rat_ID,GRO_Alpha, Study_group, hysteresis_grp) %>%
  filter(Study_group %in% c("G5","G6","G7") ) %>%
  ggplot() + geom_jitter(aes(x=hysteresis_grp, y=GRO_Alpha, col=Study_group), width = 0.1, size=5, alpha=0.5)




```


## import all physio parameters
```{r, message=FALSE, warning=FALSE}
readr::read_csv2("Data_All.csv") %>% colnames


physio = 
readr::read_csv2("Data_All.csv") %>% select(-USUBJID,-round,-dose_diet) %>%
  # select(SUBJID,diet,dose,weight_colon,length_colon,Lipocalin_T40_pg_mL,Lipocalin_T69_pg_mL,
  #        Total_distal,Total_proximal,Total_proximal,Crypt_diameter_distal,Crypt_height_distal,
  #        Goblet_cell_number_distal,Crypt_height_proximal,Crypt_diameter_proximal,Goblet_cell_number_proximal,
  #        BW_D3_diffp_D0,BW_D17_diffp_D14,BW_D31_diffp_D28,
  #        Edema_distal,Mucosa_hyperplasia_distal,Epithelial_atrophy_distal,
  #        Ulceration_distal,Cryptic_abcess_distal,Intestinal_wall_thickness_distal,
  #        Squamous_metaplasia_distal,Polyps_distal,Mononuclear_cell_infiltration_distal,Neutrophils_infiltration_distal) %>%
  filter(diet != "AIN-93G")

physio

```


```{r, fig.height=8, fig.width=10}

physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  mutate(DSS_group = ifelse(dose>=1,"High","Low")) %>%
  ggplot() +  geom_boxplot(aes(x=Study_group,y=value,fill=Study_group)) + facet_wrap(~variable, scale="free_y") + scale_fill_brewer()


# 
# 
# physio %>%
#   mutate(Rat_ID = paste0("R",SUBJID)) %>%
#   merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
#   select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
#   melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
#   na.omit() %>%
#   mutate(DSS_group = ifelse(dose>=1,"High","Low")) %>%
#   ggplot() +  
#   geom_density_ridges_gradient(aes(x=value %>% scale, y=variable , fill=..x..)) + 
#   #geom_density(aes(x=Bifidobacterium + 10^-6, fill=cat_age_val)) +
#   
#   facet_wrap(~DSS_group, scales ="free_y", ncol=2) +
#   viridis::scale_fill_viridis(option = "B") +
#   guides(fill=FALSE) +
#   scale_x_log10() + theme_dark()
# 

physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight,-dose,-Animal_id) %>%
  na.omit() %>%
  tibble::rownames_to_column("id") %>%
  select(-id,-Study_group) %>%
  tibble::column_to_rownames("Rat_ID") %>%
  dudi.pca(.,scannf=F, nf=3) %>% 
  .$li %>%
  merge(rat_metadata,by.x="row.names", by.y="Animal_id_2") %>%
  mutate(DSS_group = ifelse(Study_group == "G5" | Study_group == "G6" | Study_group == "G7", "high", "low" )) %>%
  #ggplot() + geom_histogram(aes(x=Axis1, fill=Study_group), alpha=0.5) + scale_fill_brewer() + facet_wrap(~DSS_group)
  ggplot() + geom_density(aes(x=Axis1), alpha=0.5) 
  
  

physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  mutate(DSS_group = ifelse(dose>=1,"High","Low")) %>%
  ggplot() +  geom_histogram(aes(x=value+0.1, fill=dose%>%as.character)) + facet_wrap(~variable, scale="free") + scale_fill_brewer()



physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  merge(.,hysteresis, by="Rat_ID") %>%
  #filter(dose > 0.9| dose < 0.1) %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id","hysteresis_grp")) %>%
  na.omit() %>%
  mutate(DSS_group = ifelse(dose>=1,"High","Low")) %>%
  mutate(Group = ifelse(dose>=1,hysteresis_grp%>%as.character,ifelse(dose>0,"Low DSS dose","Control\ngroup")) ) %>%
  mutate(Group = Group %>% factor(levels = c("Control\ngroup","Low DSS dose","resistant\nprofile","resilient\nprofile","no resilient\nprofile")) ) %>%
  ggplot() +  
  #geom_histogram(aes(x=value+0.1, fill=Group)) + 
  geom_boxplot(aes(y=value, fill=Group)) +
  facet_wrap(~variable, scale="free") + scale_fill_brewer() + theme_dark()

ggsave("physio_resilience.pdf")

```

test physio parameters for bimodality

fecal lipocalin, total distal score, crypt height in the distal colon, 


```{r, message=FALSE, warning=FALSE}


library(mclust)

physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  mutate(DSS_group = ifelse(dose>=1,"High","Low")) %>%
  group_by(variable) %>%
  do(data.frame(out=densityMclust(.$value, verbose=FALSE) %>% summary %>% .$G))



physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  mutate(DSS_group = ifelse(dose>=1,"High","Low")) %>%
  group_by(variable) %>%
  tidyr::nest() %>%
  mutate(model = purrr::map(data, ~ densityMclust(.$value, verbose=FALSE))) %>%
  mutate(summary = purrr::map(model,~summary)) %>%
  tidyr::unnest(model %>% purrr::map(broom::tidy)) 
  


```



```{r, fig.height=9, fig.width=12}

physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  merge(.,hysteresis, by="Rat_ID") %>%
  filter(dose > 0.9) %>%
  group_by(variable, hysteresis_grp) %>%
  summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  ungroup() %>%
  #ggplot +  geom_boxplot(aes(x=hysteresis_grp,y=value)) + facet_wrap(~variable, scale="free_y")
  ggplot(aes(x=hysteresis_grp, y=m)) +
  geom_bar(fill="white", stat="identity", position="dodge", color="black") + 
  geom_errorbar(aes(ymin=m-sem,ymax=m+sem),width=.2, position=position_dodge(.9)) +
  #scale_fill_brewer(type="qual") +
  ylab("") +
  xlab("") + facet_wrap(~variable %>% gsub("_","\n",.), scale="free_y")



physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  merge(.,hysteresis, by="Rat_ID") %>%
  filter(dose > 0.9) %>%
  #group_by(variable, hysteresis_grp) %>%
  #summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  #ungroup() %>%
  #ggplot +  geom_boxplot(aes(x=hysteresis_grp,y=value)) + facet_wrap(~variable, scale="free_y")
  group_by(hysteresis_grp, variable) %>%
  summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  ungroup() %>%
  ggplot(aes(x=hysteresis_grp, y=m)) +
  geom_bar(fill="white", stat="identity", position="dodge", color="black") + 
  geom_errorbar(aes(ymin=m-sem,ymax=m+sem),width=.2, position=position_dodge(.9)) +
  facet_wrap(~variable, scale="free_y")




akk_df %>%
  #filter(Study_group != "G2") %>%
  select(hysteresis_grp,Study_group, Rat_ID) %>% 
  unique %>%
  merge(.,histo, by="Rat_ID") %>%
  mutate(DSS_group = ifelse(Study_group == "G5" | Study_group == "G6" | Study_group == "G7", "high", "low" )) %>%
  filter(DSS_group == "high") %>%
  mutate(score_histo = as.numeric(as.character(score_histo))) %>%
  group_by(hysteresis_grp) %>%
  summarise(n = n(), m = mean(score_histo), sem = sd(score_histo)/sqrt(n())) %>%
  ungroup() %>%
  ggplot(aes(x=hysteresis_grp, y=m)) +
  geom_bar(fill="white", stat="identity", position="dodge", color="black") + 
  geom_errorbar(aes(ymin=m-sem,ymax=m+sem),width=.2, position=position_dodge(.9)) +
  #scale_fill_brewer(type="qual") +
  ylab("Histological score") +
  xlab("") +
  theme(legend.position = c(0.1, 0.9))



```


```{r, fig.height=4, fig.width=8}


physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  #merge(.,hysteresis, by="Rat_ID") %>%
  #filter(dose > 0.9) %>%
  #group_by(variable, hysteresis_grp) %>%
  #summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  #ungroup() %>%
  #ggplot +  geom_boxplot(aes(x=hysteresis_grp,y=value)) + facet_wrap(~variable, scale="free_y")
  group_by(dose, variable) %>%
  summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  ungroup() %>%
  ggplot(aes(x=dose %>% as.character, y=m)) +
  geom_bar(fill="white", stat="identity", position="dodge", color="black") + 
  geom_errorbar(aes(ymin=m-sem,ymax=m+sem),width=.1, position=position_dodge(.9)) +
  facet_wrap(~variable, scale="free_y") + xlab("DSS Dose") + ylab("parameter value")


physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  mutate(`Body_weigth_dynamics\n(3 DSS pusles)` = BW_D3_diffp_D0 + BW_D17_diffp_D14 + BW_D31_diffp_D28) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  filter(variable %in% c("length_colon","Total_distal","Body_weigth_dynamics\n(3 DSS pusles)")) %>%
  na.omit() %>%
  #merge(.,hysteresis, by="Rat_ID") %>%
  #filter(dose > 0.9) %>%
  #group_by(variable, hysteresis_grp) %>%
  #summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  #ungroup() %>%
  #ggplot +  geom_boxplot(aes(x=hysteresis_grp,y=value)) + facet_wrap(~variable, scale="free_y")
  group_by(dose, variable) %>%
  summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  ungroup() %>%
  ggplot(aes(x=dose %>% as.character, y=m)) +
  geom_bar(fill="white", stat="identity", position="dodge", color="black") + 
  geom_errorbar(aes(ymin=m-sem,ymax=m+sem),width=.1, position=position_dodge(.9)) +
  facet_wrap(~variable, scale="free_y") + xlab("DSS Dose") + ylab("parameter value")

physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  mutate(`Body_weigth_dynamics` = BW_D3_diffp_D0 + BW_D17_diffp_D14 + BW_D31_diffp_D28) %>%
  rename(`2 colon length\n(cm)` = length_colon, 
         `3 distal\nhistological score` = Total_distal, 
         `1 Body weight variations\nduring the 3 DSS pulses\n(gram)` = Body_weigth_dynamics) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  filter(variable %in% c("2 colon length\n(cm)","3 distal\nhistological score","1 Body weight variations\nduring the 3 DSS pulses\n(gram)")) %>%
  mutate(variable = variable %>% as.character) %>%
  #filter(variable %in% c("length_colon","Total_distal","Body_weigth_dynamics\n(3 DSS pusles)")) %>%
  na.omit() %>%
  #merge(.,hysteresis, by="Rat_ID") %>%
  #filter(dose > 0.9) %>%
  #group_by(variable, hysteresis_grp) %>%
  #summarise(n = n(), m = mean(value), sem = sd(value)/sqrt(n())) %>%
  #ungroup() %>%
  #ggplot +  geom_boxplot(aes(x=hysteresis_grp,y=value)) + facet_wrap(~variable, scale="free_y")
  #group_by(dose, variable) %>%
  ggplot(aes(x=dose %>% as.character, y=value)) +
  geom_boxplot(aes(fill=dose%>% as.character)) + 
  geom_jitter(aes(fill=dose%>% as.character), alpha=0.5, width = 0.1) +
  scale_fill_manual(values = Grp_color[-1]) + guides(fill=FALSE) +
  facet_wrap(~variable %>% stringr::str_sub(start=2) , scale="free_y") + xlab("DSS Dose (%)") + ylab("parameter value") -> p_physio

save(p_physio, file="p_physio.rda")

p_physio


```


analyse interclass
```{r}


physio %>%
  mutate(Rat_ID = paste0("R",SUBJID)) %>%
  merge(.,rat_metadata,by.x="Rat_ID",by.y="Animal_id_2") %>%
  select(-SUBJID,-diet,-Cage_provider,-Animal.Identification,-Filiation,-Bacteroides,-Weight) %>%
  melt(id.vars=c("Rat_ID","dose","Study_group","Animal_id")) %>%
  na.omit() %>%
  merge(.,hysteresis, by="Rat_ID") %>%
  filter(dose > 0.9) %>%
  dcast(Rat_ID+hysteresis_grp~variable, value.var="value") -> df

df_pca = ade4::dudi.pca(df[-c(1,2)] %>% na.omit, scannf=F, nf=10)

#bca(df_pca, fac = df %>% na.omit %>% .$hysteresis_grp, scannf=F, nf=2) %>% randtest()


cbind(df %>% na.omit,df_pca$li) %>%
  ggplot() + geom_point(aes(x=Axis1,y=Axis2,col=hysteresis_grp))




```


